---
title: "Lab 03 - Nobel laureates"
author: "Lindsay Stall"
date: "01/20/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{rview}
view(nobel)
```


There are 935 observations, and 26 variables. Each row represents information of a different nobel prize winning scientist.

### Exercise 2

```{r}
nobel_living <- nobel

nobel_living %>% 

  filter(country != "NA") %>%

  filter(gender != "org") %>%

  filter(is.na(died_date))
```


### Exercise 3

```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r}
nobel_living_sciences <- filter(nobel_living_science, country_us != "NA")
```



```{r}
ggplot(nobel_living_sciences, aes(y = country_us))+
geom_bar()+
facet_wrap(~category, ncol = 1)
```
While the US has many more nobel prize winning scientists than other countries in the field of economics, it has fewer in the field of Chemistry and comparable amounts in Medicine and Physics. 


### Exercise 4
```{r}
nobel_living <- nobel_living %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living <- nobel_living %>%
  select(born_country_us)
summary(as.factor(nobel_living$born_country_us))
```
 271 were born in the US.

### Exercise 5

```{r}
ggplot(nobel_living_sciences, aes(y = country_us), fill(born_country_us))+
geom_bar()+
facet_wrap(~category, ncol = 1)

```


### Exercise 6

```{r}

```

