---
title: "Lab 03 - Nobel laureates"
author: "Lindsay Stall"
date: "01/20/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r summ}
summary(nobel)
```


There are 935 observations, and 26 variables. Each row represents information of a different nobel prize winning scientist.

### Exercise 2

```{r filter}
nobel_living <- nobel

nobel_living %>% 

  filter(country != "NA") %>%

  filter(gender != "org") %>%

  filter(is.na(died_date))
```


### Exercise 3

```{r facet}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))%>%
  filter(country_us != "NA")
```


```{r}
ggplot(nobel_living_science, aes(y = country_us))+
geom_bar()+
facet_wrap(~category, ncol = 1)
```
While the US has many more nobel prize winning scientists than other countries in the field of economics, it has fewer in the field of Chemistry and comparable amounts in Medicine and Physics. 


### Exercise 4
```{r born-country}
nobel_living <- nobel_living %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other"))
```

```{r}
nobel_living %>%
  select(born_country_us)
summary(as.factor(nobel_living$born_country_us))
```
 271 were born in the US.

### Exercise 5

```{r mutate}
nobel_living_science <- nobel_living %>%
  mutate(born_country_us = if_else(born_country == "USA", "USA", "Other"))
```

```{r filtering}
nobel_living_science <- nobel_living %>% 
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))%>%
  filter(country_us != "NA")
```


```{r graphs}
ggplot(nobel_living_science, aes(x = country_us, color = born_country_us, fill= born_country_us))+
geom_bar()+
facet_wrap(~category, ncol = 1)

```

Buzzfeed said that most living laureates  were based in the US when they were born. The above graph shows that that is only true for economics. While there are a lot of people born outside of the US who got the award, of the ones who received it in US, most were born in the US.

### Exercise 6

```{r}
nobel_living_science %>%
  filter(country_us == "USA") %>%
  filter(born_country_us == "Other") %>%
  count(born_country) %>%
  arrange(desc(n))
```
Most of the winners from the US were born in the UK.
